---
title: 'BA Homework #5'
author: "Kunyang Que"
date: "10/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem I: Bicycling World

## 1.1 Import Data

```{r}
data_bicycle<-data.frame(
  Model=c("Fierro 7B", "HX 5000", "Durbin Ultralight", "Schmidt", "WSilton Advanced", "bicyclette v€lo", "Supremo Team", "XTC Racer", "D'Onofrio Pro", "Americana #6"), 
  Weight=c(17.9, 16.2, 15.0, 16.0, 17.3, 13.2, 16.3, 17.2, 17.7, 14.2),
  Price=c(2200, 6350, 8470, 6300, 4100, 8700, 6100, 2680, 3500, 8100))
```

## 1.2 Scatter Chart

```{r}
plot(data_bicycle$Weight, data_bicycle$Price)
```

The Price decreases as Weight increases

## 1.3 Regression

```{r}
x<-data_bicycle$Weight
y<-data_bicycle$Price

model_bicycle<-lm(y ~ x)
model_bicycle
summary(model_bicycle)
```

## 1.4 Interpretations

When the weight is raised by 1 level, the price will drop by $1439.

P-value:
(Intercept)  2.15e-05 
x            9.99e-05 

If we set the significant level as 0.05, the p-value for both parameters is less than the signigicant level, which means I should reject null hypothesis that both regression parameters are equal to 0.

## 1.5 Variation

R-squared:  0.8637
86.37% of the data fit the regression model.

# Problem II: Assembly Line Problem

## 2.1 Import Data

```{r}
data_line<-data.frame(
  Line_Speed=c(20, 20, 40, 30, 60, 40),
  Number=c(21, 19, 15, 16, 14, 17))
```

## 2.2 Scatter Chart

```{r}
plot(data_line$Line_Speed, data_line$Number)
```

The number of defective parts found decreases when the line speed increases.

## 2.3 Regression

```{r}
x<-data_line$Line_Speed
y<-data_line$Number

model_line<-lm(y ~ x)
model_line
summary(model_line)
```

## 2.4 Significance Level

* The p-value for B0 is less than significance level 0.01. Therefore, I reject null hypothesis 
that B0 is equal to 0 at 0.01 level of significance. 

* The p-value for B1 is greater than significance level 0.01.For B1. Therefore, I retain the null hypothesis that B1 is equal to 0 at 0.01 level of significance.

## 2.5 Variation

R-squared:  0.7391
73.91% of the data fit the regression model.

# Problem III: Jensen Tire & Auto Problem

## 3.1 Import Data

```{r}
data_jensen<-data.frame(
  weekly_usage=c(13, 10, 20, 28, 32, 17, 24, 31, 40, 38),
  maintenance=c(17.0, 22.0, 30.0, 37.0, 47.0, 30.5, 32.5, 39.0, 51.5, 40.0))
```

## 3.2 Scatter Chart

```{r}
plot(data_jensen$weekly_usage, data_jensen$maintenance)
```

The scatter chart shows that there is a strong linear relationship between the weekly usage and the maintenance. With the weekly usage increase, the maintenance increase as well.

## 3.3 Regression

```{r}
x<-data_jensen$weekly_usage
y<-data_jensen$maintenance

model_jensen<-lm(y ~ x)
model_jensen
summary(model_jensen)
```

# Problem IV: Toyota Problem

## 4.1 Import Data

```{r}
data_toyota<-data.frame(
  miles=c(22, 29, 36, 47, 63, 77, 73, 87, 92, 101, 110, 28, 59, 68, 68, 91, 42, 65, 110),
  price=c(16.2, 16.0, 13.8, 11.5, 12.5, 12.9, 11.2, 13.0, 11.8, 10.8, 8.3, 12.5, 11.1, 15.0, 12.2, 13.0, 15.6, 12.7, 8.3))
```

## 4.2 Scatter Chart

```{r}
plot(data_toyota$miles, data_toyota$price)
```

## 4.3 Regession Model

```{r}
x<-data_toyota$miles
y<-data_toyota$price

model_toyota<-lm(y ~ x)
model_toyota
summary(model_toyota)
```
## 4.4 Significance Test

If we set the significance level as 0.01, the summary of the model shows that the p-value for both regression parameters are less than the significance level.Therefore, I reject null hypothesis that both regression parameters are equal to 0 at 0.01 level of significance.

## 4.5 Variation

Multiple R-squared:  0.5387
53.87% of the data fit the regression model.

## 4.6 Predict Price and Residual

```{r}
prediction<-function(x) {
  return (16.46976 - 0.05877 * x) 
}

pre_price<-prediction(data_toyota$miles)
pre_price
```

```{r}
residual<-data_toyota$price - pre_price
residual
```

Miles: 47000  Price: $11500
Miles: 28000  Price: $12500

## 4.7 Prediction

```{r}
prediction(60)
```

According to the model, I should have priced it at $12,944. But before I put a price on the seller, I will estimate about other aspects of the car.

# Problem V: Dodger Stadium Attendance

## 5.1 Import Data

```{r}
url<-"https://raw.githubusercontent.com/jcbonilla/BusinessAnalytics/master/BAData/dodgers.csv"
data_dodger<-read.csv(url, stringsAsFactors = TRUE, header = T)
head(data_dodger)
```

## 5.2 Exploratory Data Analysis

### a.
```{r}
summary(data_dodger)
```

Cap: 2
Shirt: 3
Fireworks: 14
Bobblehead: 11
Total: 30

### b. 

```{r}
library(dplyr)
data_dodger<-mutate(data_dodger, promotion=if_else(data_dodger$cap == "YES", 1, 
                                                   if_else(data_dodger$shirt == "YES", 1, 
                                                           if_else(data_dodger$fireworks == "YES", 1, 
                                                                   if_else(data_dodger$bobblehead == "YES", 1, 0)))))
```

```{r}
data_pro<-subset(data_dodger, data_dodger$promotion == 1)
data_non<-subset(data_dodger, data_dodger$promotion == 0)

mean(data_pro$attend)
mean(data_non$attend)
```

### c.

Temperature
```{r}
library(ggplot2)
ggplot(data_dodger, aes(x=temp, y=attend)) + geom_point()
```


### d. 

```{r}
library(sqldf)
mean(data_dodger$attend)
```

```{r}
dodger_team<-sqldf("select opponent, avg(attend) as avg_attend from data_dodger group by opponent order by avg_attend desc")
dodger_team
```

```{r}
sqldf("select opponent, avg_attend from dodger_team where avg_attend > 41040.07")
```

## 5.3 Predictive Modeling

### a. 

```{r}
sqldf("select avg(attend) from data_dodger where bobblehead == 'YES'")
```

### b.

```{r}
sqldf("select avg(attend) from data_dodger where cap == 'YES'")
```

```{r}
sqldf("select avg(attend) from data_dodger where shirt == 'YES'")
```

```{r}
sqldf("select avg(attend) from data_dodger where fireworks == 'YES'")
```

When Bobblehead had promotions, the attendance was the highest of all times. Therefore, I would say the Bobblehead promotions are better than all other promotions.

### c.

```{r}
bobble<-sqldf("select * from data_dodger where bobblehead == 'YES'")
bobble
```
```{r}
mean(bobble$attend)
mean(data_non$attend)
mean(bobble$attend) - mean(data_non$attend)
```

Compared with the day without any promotion, each Bobblehead promotion brings an increase in attendance of 14943.56.

```{r}
summer<-sqldf("select * from data_dodger where month in ('JUN', 'JUL', 'AUG')")
summer
```

```{r}
length(summer$attend)
sum(summer$attend)
```

Now we have 36 events with the attendance of 1599348.

```{r}
(1599348 - 38201.08 * 36) / 14943.56
```
In order to reach an attendance of 1599348, Bobblehead Promotion shall be held at least 15 times without consideration of other promotions.


