---
title: "BA Mid Project Deep Analysis"
author: "Group 5: Gabriela Caballero, Kunyang Que, Ullas Srivastava, Yangxin Liu"
date: "10/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Data Preprocessing in MySQL

## 1.1 Delete columns that are all null

```sql
alter table tb_mid_rivier_users drop column citizenship_country;
alter table tb_mid_rivier_users drop column lives_in_us;
alter table tb_mid_rivier_users drop column birth_place;
alter table tb_mid_rivier_users drop column permanent_resident_number;
alter table tb_mid_rivier_users drop column intended_school;
alter table tb_mid_rivier_users drop column intended_concentration;
alter table tb_mid_rivier_users drop column has_24credits;
alter table tb_mid_rivier_users drop column has_transfer_credits;
alter table tb_mid_rivier_users drop column hs_course_credits;
alter table tb_mid_rivier_users drop column hs_course_title;
alter table tb_mid_rivier_users drop column hs_degree_earned;
alter table tb_mid_rivier_users drop column hs_enrollment_date;
alter table tb_mid_rivier_users drop column hs_major;
alter table tb_mid_rivier_users drop column college_course_credits;
alter table tb_mid_rivier_users drop column college_course_title;
alter table tb_mid_rivier_users drop column college_enrollment_date;
alter table tb_mid_rivier_users drop column college_street2;
alter table tb_mid_rivier_users drop column address_mailing_eps_code;
alter table tb_mid_rivier_users drop column address_mailing_lat;
alter table tb_mid_rivier_users drop column address_mailing_lng;
alter table tb_mid_rivier_users drop column address_mailing_ismailing;
alter table tb_mid_rivier_users drop column milestone_prospect_source;
alter table tb_mid_rivier_users drop column milestone_checklist_complete_date;
alter table tb_mid_rivier_users drop column milestone_application_decision_date;
alter table tb_mid_rivier_users drop column milestone_waitlist_date;
alter table tb_mid_rivier_users drop column milestone_visit_date;
alter table tb_mid_rivier_users drop column milestone_interview_date;
alter table tb_mid_rivier_users drop column first_source_utm_content;
alter table tb_mid_rivier_users drop column first_source_utm_term;
alter table tb_mid_rivier_users drop column web_source_utm_content;
alter table tb_mid_rivier_users drop column web_source_utm_term;
alter table tb_mid_rivier_users drop column app_source_app_name;
alter table tb_mid_rivier_users drop column app_source_utm_content;
alter table tb_mid_rivier_users drop column app_source_utm_term;
alter table tb_mid_rivier_users drop column intended_concentration_name;
alter table tb_mid_rivier_users drop column intended_school_name;
alter table tb_mid_rivier_users drop column milestone_conditional_admit_date;
alter table tb_mid_rivier_users drop column milestone_hold_date;
alter table tb_mid_rivier_users drop column territory;
alter table tb_mid_rivier_users drop column profile_type;
```

## 1.2 Create analysis tables

### 1.2.1 Basic infomation

```sql
create table tb_mid_ana_basic(
    select element_id,
           gender,
           date_of_birth,
           labels,
           citizenship_status,
           country_of_birth,
           english_native_language,
           years_lived_in_us,
           native_language,
           us_visa_holder,
           visa_type,
           visa_required
    from tb_mid_rivier_users
);
```

### 1.2.2 Student Information

```sql
create table tb_mid_ana_stu_info(
    select element_id,
           applying_f1,
           intended_major,
           student_type,
           intended_term,
           intended_campus,
           academic_load,
           applied_before,
           veteran_funds
    from tb_mid_rivier_users
);
```

### 1.2.3 House Infomation

```sql
create table tb_mid_ana_house_info(
    select element_id,
           housing,
           household_income,
           household_income_median,
           household_persons_count,
           household_value
    from tb_mid_rivier_users
);
```

### 1.2.4 High school and college

```sql
create table tb_mid_ana_school(
    select element_id,
           hs_ceeb,
           hs_name,
           hs_end_date,
           hs_gpa,
           hs_graduate_date,
           hs_graduated,
           hs_street1,
           hs_street2,
           hs_city,
           hs_state,
           hs_country,
           hs_zipcode,
           college_ceeb,
           college_name,
           college_degree_earned,
           college_end_date,
           college_gpa,
           college_graduate_date,
           college_graduated,
           college_major,
           college_street1,
           college_city,
           college_state,
           college_country,
           college_zipcode
    from tb_mid_rivier_users
);
```

### 1.2.5 Address Infomation

```sql
create table tb_mid_ana_address(
    select element_id,
           address_home_city,
           address_home_country,
           address_home_state,
           address_home_province,
           address_home_zipcode,
           address_home_zipcode4,
           address_home_county,
           address_home_eps_code,
           address_home_lat,
           address_home_lng,
           address_home_ismailing,
           address_home_minutes_to_campus,
           address_home_miles_to_campus,
           address_mailing_city,
           address_mailing_country,
           address_mailing_state,
           address_mailing_province,
           address_mailing_zipcode,
           address_mailing_zipcode4,
           address_mailing_county,
           address_mailing_minutes_to_campus,
           address_mailing_miles_to_campus
    from tb_mid_rivier_users
);
```

### 1.2.6 Milestone infomation

```sql
create table tb_mid_ana_milestone(
    select element_id,
           milestone_created_date,
           milestone_prospect_date,
           milestone_application_start_date,
           milestone_application_start_guid,
           milestone_application_start_apptype,
           milestone_application_start_term,
           milestone_application_complete_date,
           milestone_application_complete_guid,
           milestone_application_complete_apptype,
           milestone_application_complete_term,
           milestone_application_submit_date,
           milestone_application_submit_guid,
           milestone_application_submit_apptype,
           milestone_application_submit_term,
           milestone_admit_date,
           milestone_admit_term,
           milestone_deposit_date,
           milestone_deposit_term,
           milestone_withdraw_date,
           milestone_withdraw_term,
           milestone_unsubscribe_sms_date,
           milestone_unsubscribe_sms_number,
           milestone_unsubscribe_email_date,
           milestone_unsubscribe_sms_email,
           milestone_denied_date,
           milestone_enroll_date,
           milestone_hardbounce_sms_date,
           milestone_hardbounce_email_date
    from tb_mid_rivier_users
);
```

# 2. Load Data and Import Library

## 2.1 Import Library

```{r}
library(dplyr)
library(sqldf)
library(ggplot2)
library(OpenStreetMap)
```

## 2.2 Load Data

```{r}
data_basic<-read.csv("/Users/kunyangque/Documents/NYUTandon/course/Business Analysis/Midterm Project/data/01_tb_mid_ana_basic.csv", header=T, na.strings=c("","NA"))
data_stu_info<-read.csv("/Users/kunyangque/Documents/NYUTandon/course/Business Analysis/Midterm Project/data/02_tb_mid_ana_stu_info.csv", header=T, na.strings=c("","NA"))
schools<-read.csv("/Users/kunyangque/Documents/NYUTandon/course/Business Analysis/Midterm Project/data/03_tb_mid_ana_school.csv", header = TRUE, na.strings=c("","NA"))
data_housing<-read.csv("/Users/kunyangque/Documents/NYUTandon/course/Business Analysis/Midterm Project/data/04_tb_mid_ana_address.csv")
milestones<-read.csv("/Users/kunyangque/Documents/NYUTandon/course/Business Analysis/Midterm Project/data/05_tb_mid_ana_milestone.csv", header=T, na.strings=c("","NA"))

data_major_list<-read.csv("/Users/kunyangque/Documents/NYUTandon/course/Business Analysis/Midterm Project/data/tb_major_list.csv", header=T, na.strings=c("","NA"))
```

## 3. Characteristics from Basic Information

### 3.1 Age

Select students between the ages of **15** and **24**

```{r}
Year<-format(as.Date(data_basic$date_of_birth, format = "%m/%d/%Y"), "%Y")
Year<-as.integer(Year)

data_basic<-mutate(data_basic, birth_year=if_else(Year<20, Year+2000, Year+1900))
data_basic$age<-2020 - data_basic$birth_year

db_student<-sqldf("select * from data_basic where age >= 15 and age <= 24")

ggplot(data = db_student, mapping = aes(x=age, fill=age)) + geom_bar(stat="count", width=0.5) + 
geom_text(stat='count',aes(label=..count..), vjust=-0.7, size=3.5)
```

### 3.2 Major

The **top 10** most popular majors

```{r}
stu_info<-inner_join(db_student, data_stu_info, by = "element_id")
major<-sqldf("select intended_major as guid, count(*) as num from stu_info group by intended_major order by num desc")

join_major<-inner_join(major, data_major_list, by="guid")
major_top_10<-sqldf("select major_name, num from join_major order by num desc limit 10")
ggplot(data = major_top_10, mapping = aes(x=major_name, y=num)) + 
geom_bar(stat="identity", width=0.8) + 
geom_text(aes(label=num), vjust=-0.7, size=3.5)
```

# 4. Characteristics from Milestone

```{r}
join_milestone<-inner_join(db_student, milestones, by="element_id")
```

## 4.1 Applications started by month
```{r}
join_milestone$milestone_application_start_date<-as.Date(join_milestone$milestone_application_start_date,"%m/%d/%y")

#Extracting months from the start application date
join_milestone$createdMonth<-months(join_milestone$milestone_application_start_date)

#Counting applications started by month
applCreatedByMonth<-aggregate(na.omit(join_milestone$element_id),by=list(join_milestone$createdMonth),FUN="length")
applCreatedByMonth

#Ordering 
applCreatedByMonth$Group.1 <- factor(applCreatedByMonth$Group.1,levels = c("January", "February", "March", "April","May", "June","July","August","September","October","November","December"))
```

```{r}
#Ploting
ggplot(applCreatedByMonth,aes(x=Group.1,y=x)) +
geom_bar(stat="identity",width=0.5) +
geom_text(aes(label=x), vjust=-0.7, size=3.5) +
ggtitle("Applications Started by Month") + 
ylim(0,max(applCreatedByMonth$x))
```

## 4.2 AVG time to complete an application in months
```{r cars}
#Converting into date format
join_milestone$milestone_application_complete_date<-as.Date(join_milestone$milestone_application_complete_date,"%m/%d/%y")
join_milestone$milestone_application_start_date<-as.Date(join_milestone$milestone_application_start_date,"%m/%d/%y")
join_milestone$daysToCompleteApl<-as.numeric(join_milestone$milestone_application_complete_date - join_milestone$milestone_application_start_date)
#Average days to complete an application
summary(join_milestone$daysToCompleteApl)
```

```{r}
boxplot(join_milestone$daysToCompleteApl)
```

## 4.3 Percentage of applications started that end up being submited 

```{r}
appStarted=subset(join_milestone,!is.na(join_milestone$milestone_application_start_date))
appSubmited=subset(join_milestone,!is.na(join_milestone$milestone_application_submit_date))
appDiscarded=length(appStarted$element_id)-length(appSubmited$element_id)
#PercApSubmited=(length(appSubmited$element_id)*100)/length(appStarted$element_id)
#PercApDiscarded=appStarted- PercApSubmited
slices <- c(length(appSubmited$element_id), appDiscarded)
lbls<-c("Applications submited","Applications discarded")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
pieChart=pie(slices,labels = lbls, col=rainbow(length(lbls)),
   main="Pie Chart of Applications submitted")
```

## 4.4 Applications submitted that are admited
```{r}
appsAdmited=subset(join_milestone,!is.na(join_milestone$milestone_admit_date))
appsDenied=subset(join_milestone,!is.na(join_milestone$milestone_denied_date))
PercApAdmited=length(appsAdmited$element_id)
PercApDenied=length(appsDenied$element_id)
PercpApsUnk=length(appStarted$element_id)-(PercApAdmited+PercApDenied)
slices2 <- c(PercApAdmited, PercApDenied,PercpApsUnk)
lbls2<-c("Applications Admited","Applications Denied","Applications without information")
pct2 <- round(slices2/sum(slices2)*100)
lbls2 <- paste(lbls2, pct2) # add percents to labels
lbls2 <- paste(lbls2,"%",sep="") # ad % to labels
pieChart2=pie(slices2,labels = lbls2, col=rainbow(length(lbls2)),
   main="Pie Chart of Admited and Denied Applications")
```
## 4.5 Applications admitted per month
```{r}
join_milestone$milestone_admit_date<-as.Date(join_milestone$milestone_admit_date,"%m/%d/%y")
#Extracting months from the admitions application date
join_milestone$admitedMonth<-months(join_milestone$milestone_admit_date)
#Counting applications admited by month

applAdmitedByMonth<-aggregate(na.omit(join_milestone$element_id),by=list(join_milestone$admitedMonth),FUN="length")
applAdmitedByMonth
#Ordering 
applAdmitedByMonth$Group.1 <- factor(applAdmitedByMonth$Group.1,levels = c("January", "February", "March", "April","May", "June","July","August","September","October","November","December"))
```

```{r}
#Ploting
ggplot(applAdmitedByMonth,aes(x=Group.1,y=x)) +
geom_bar(stat="identity",width=0.5) +
ggtitle("Applications Admited by Month") + 
geom_text(aes(label=x), vjust=-0.7, size=3.5) + 
ylim(0,max(applAdmitedByMonth$x))
```

## 4.6 Applicants admited that end up enrolling and withdrawing
```{r}
#Converting into date format
appsEnrolled=subset(join_milestone,!is.na(join_milestone$milestone_enroll_date))
PercApEnrolled=length(appsEnrolled$element_id)
appsWithdraw=subset(join_milestone,!is.na(join_milestone$milestone_withdraw_date))
PercApWithdrawed=length(appsWithdraw$element_id)
PerceApAdmitUnkonw=length(appsAdmited$element_id)-(PercApEnrolled+PercApWithdrawed)
PercApEnrolled
PercApWithdrawed
PerceApAdmitUnkonw
```

```{r}
slices3 <- c(PercApEnrolled, PercApWithdrawed, PerceApAdmitUnkonw)
lbls3<-c("Applicants Enrolled","Applications Withdrawed", "Applications without info")
pct3 <- round(slices3/sum(slices3)*100)
lbls3 <- paste(lbls3, pct3) # add percents to labels
lbls3 <- paste(lbls3,"%",sep="") # ad % to labels
pieChart3=pie(slices3,labels = lbls3, col=rainbow(length(lbls3)),
   main="Pie Chart of Enrolled and Wirdrawed Applications")
```

## 4.7 Enrollings per Month
```{r}
join_milestone$milestone_enroll_date<-as.Date(join_milestone$milestone_enroll_date,"%m/%d/%y")
#Extracting months from the enrolled application date
join_milestone$enrollMonth<-months(join_milestone$milestone_enroll_date)
#Counting applications enrolled by month

applEnrolledByMonth<-aggregate(na.omit(join_milestone$element_id),by=list(join_milestone$enrollMonth),FUN="length")
applEnrolledByMonth
#Ordering 
applEnrolledByMonth$Group.1 <- factor(applEnrolledByMonth$Group.1,levels = c("January", "February", "March", "April","May", "June","July","August","September","October","November","December"))
```

```{r}
#Plotting
ggplot(applEnrolledByMonth,aes(x=Group.1,y=x)) +
geom_bar(stat="identity",width=0.5) +
ggtitle("Applications Enrolled by Month") + 
geom_text(aes(label=x), vjust=-0.7, size=3.5) + 
ylim(0,max(applEnrolledByMonth$x))
```

# 5. Characteristics from school infomation

```{r}
join_schools<-inner_join(db_student, schools, by="element_id")
```

## 5.1 High School GPA

```{r}
library(ggplot2)

schools$hs_gpa <- as.integer(schools$hs_gpa)
typeof(schools$hs_gpa)
summary(schools$hs_gpa)

high_gpa <- aggregate(na.omit(schools$element_id), by = list(schools$hs_gpa), FUN = "length")
high_gpa

high_gpa<- subset(high_gpa,high_gpa$x<4 & high_gpa$x>0)

ggplot(high_gpa,aes(x = reorder (Group.1, -x), y=x) ) +
geom_bar(stat="identity",width=0.5) + theme(axis.text.x = element_text(angle = 90)) + ggtitle("Students by High School State") + ylim(0,max(high_gpa$x))

```

## 5.2 High School State

```{r}
hs_state <- aggregate(na.omit(join_schools$element_id), by = list(join_schools$hs_state), FUN = "length")
hs_state<- subset(hs_state,hs_state$x>1000)

ggplot(hs_state,aes(x = reorder (Group.1, -x), y=x) ) +
geom_bar(stat="identity",width=0.5) + 
theme(axis.text.x = element_text(angle = 90)) + 
geom_text(aes(label=x), vjust=-0.7, size=3.5) +
ggtitle("Students by High School State") + 
ylim(0,max(hs_state$x))
```

## 5.3 College Name 

```{r}
college_name <- aggregate(na.omit(schools$element_id), by = list(schools$college_name), FUN = "length")

college_name<- subset(college_name,college_name$x>100)

ggplot(college_name,aes(x = reorder (Group.1, -x), y=x) )+
geom_bar(stat="identity",width=0.8) + theme(axis.text.x = element_text(angle = 90)) + ggtitle("Students by High School State") + ylim(0,max(college_name$x))
```

## 5.4 College State

```{r}
college_state <- aggregate(na.omit(join_schools$element_id), by = list(join_schools$college_state), FUN = "length")

college_state<- subset(college_state,college_state$x>20)

ggplot(college_state,aes(x = reorder (Group.1, -x), y=x) ) +
geom_bar(stat="identity",width=0.8) + theme(axis.text.x = element_text(angle = 90)) + ggtitle("Students by High School State") + ylim(0,max(college_state$x))
```

## 5.5 College GPA

```{r}
college_gpa <- join_schools$college_gpa
summary(college_gpa)

hist(as.numeric(college_gpa))
```

## 5.6 College Graduation

```{r}
Graduate_C <- subset(join_schools, join_schools$college_graduated == "t")
Failed_C <- subset(join_schools, join_schools$college_graduated == "f")

Graduate_C_length <- length(Graduate_C$element_id)
Failed_C_length <- length(Failed_C$element_id)

graduation <- c(Graduate_C_length, Failed_C_length)
grad_labels <- c("Graduation","Failed")

portion <- round(graduation/sum(graduation)*100)
pic_pie <- paste(grad_labels, portion)
pic_pie <- paste(grad_labels,"%",sep="")

pie(graduation, labels = grad_labels,
     main="Pie Chart of College Graduation")

```

# 6. Characteristics from Housing Information

```{r}
join_housing<-inner_join(db_student, data_housing, by="element_id")
```

## 6.1 Mapping the data_housing for all the applicant's Countries
```{r}
#usmap=openmap(
#c(round(max(data_housing$address_home_lat,na.rm=T),2),round(min(data_housing$address_home_lng,na.rm=T),2)),
#c(round(min(data_housing$address_home_lat,na.rm=T),2),round(max(data_housing$address_home_lng,na.rm=T),2)),
#type="osm")

```
```{r}
#autoplot.OpenStreetMap(openproj(usmap))+geom_point(data_housing = data_housing,mapping = aes(x=address_home_lng,y=address_home_lat))
```

## 6.2 Histogram of the most popular cities

```{r}
#addr-city
temp<-select(join_housing,col="address_home_city")
temp<-subset(temp,!is.na(temp) & temp != "" &temp!=0)
temp<-aggregate(x=temp,by = list(temp$col),FUN=length)

#filtering by the number of col
temp<- subset(temp,temp$col>1500)

ggplot(temp)+geom_bar(aes(x=reorder(Group.1,-col),y=col),stat="identity")+ theme(axis.text.x = element_text(angle = 90))+
ggtitle("Most Popular City")

```

## 6.3 Histogram of the most popular States

``` {R}
#addr-state
temp<-select(join_housing,col="address_home_state")
temp<- subset(temp,!is.na(temp) & temp != "" &temp!=0)
temp<-aggregate(x=temp,by = list(temp$col),FUN=length)

temp<- subset(temp,temp$col>100)
ggplot(temp)+geom_bar(aes(x=reorder(Group.1,-col),y=col),stat="identity")+ theme(axis.text.x = element_text(angle = 90))+
ggtitle("Most Popular State")
```

## 6.4 Histogram of the most popular Counties

``` {R}
#addr-county
temp<-select(join_housing,col="address_home_county")
temp<-subset(temp,!is.na(temp) & temp != "" &temp!=0)
temp<-aggregate(x=temp,by = list(temp$col),FUN=length)

#clean
temp<- subset(temp,temp$col>10000)
ggplot(temp)+geom_bar(aes(x=reorder(Group.1,-col),y=col),stat="identity")+ theme(axis.text.x = element_text(angle = 90)) +
ggtitle("Most Popular County")
```
