---
title: "BA Assignment 6"
author: "Kunyang Que"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1: Progresso Soup Sales

## 1.1 Load Data

```{r}
url<-"https://raw.githubusercontent.com/jcbonilla/BusinessAnalytics/master/BAData/Progresso_Soup.csv"
data_progresso<-read.csv(url, header = T, stringsAsFactors = F)
head(data_progresso)
```

## 1.2 Dummy Variable

```{r}
library(dplyr)
data_progresso<-mutate(data_progresso, is_winter=if_else(data_progresso$Month >= 10 | data_progresso$Month <= 2, 1 ,0))
head(data_progresso)
```

```{r}
library(sqldf)

data_by_month<-sqldf("select Month, sum([Sales.Progresso]) as total_sales_progresso, sum([Category_Sales]) as total_category_sales, is_winter from data_progresso group by Month")
data_by_month
```

```{r}
library(ggplot2)
ggplot(data = data_by_month, mapping = aes(x = Month, y = total_sales_progresso)) + geom_bar(stat = "identity", width = 0.5)
```

```{r}
library(ggplot2)
ggplot(data = data_by_month, mapping = aes(x = Month, y = total_category_sales)) + geom_bar(stat = "identity", width = 0.5)
```

* As we can see from the bar chart, the sales of soup vary from season to season. Sales are high in autumn and winter and low in spring and summer.

```{r}
sqldf("select Month, total_sales_progresso/total_category_sales as market_share, is_winter from data_by_month")
```

```{r}
sqldf("select is_winter, sum(total_sales_progresso)/sum(total_category_sales) as market_share from data_by_month group by is_winter")
```

## 1.3 Linear Regression

```{r}
y<-data_progresso$Sales.Progresso
x1<-data_progresso$Month
x2<-data_progresso$Price.Campbell
x3<-data_progresso$Price.PL
x4<-data_progresso$Price.Progresso
x5<-data_progresso$Category_Sales
x6<-data_progresso$is_winter

model_progresso<-lm(y ~ x1 + x2 + x3 + x4 + x5)
model_progresso
summary(model_progresso)
```

* If we set the significance level as 0.01, the summary of the model shows that the p-value for all regression parameters are less than the significance level.Therefore, I reject null hypothesis that all the regression parameters are equal to 0 at 0.01 level of significance.

## 1.4 Validate

```{r}
predict_function<-function(x1, x2, x3, x4, x5) {
  return (-1467.1228 - 23.8122 * x1 + 2223.6653 * x2 + 1139.3441 * x3 - 1942.8330 * x4 + 0.3134 * x5)
}

predict_value<-predict_function(x1, x2, x3, x4, x5)
cost<-sum((predict_value - y) ** 2) / length(y)
cost
```

* The R-squared of the model is 0.7138. 71.38% of the data fit the regression model.

# Problem 2: Diamond Quotes

## 2.1 Load Data

```{r}
url<-"https://raw.githubusercontent.com/jcbonilla/BusinessAnalytics/master/BAData/Diamonds.csv"
data_diamond<-read.csv(url, header = TRUE, stringsAsFactors = TRUE)
head(data_diamond)
```

```{r}
summary(data_diamond)
```

```{r}
column_convert<-sapply(data_diamond, is.factor)
convert_data<-sapply(data_diamond[, column_convert], unclass)
data_diamond_numeric<-cbind(data_diamond[, !column_convert], convert_data)
head(data_diamond_numeric)
```

## 2.2 Model1

```{r}
y<-data_diamond_numeric$Price
x1<-data_diamond_numeric$Carat
x2<-data_diamond_numeric$Wholesaler
x3<-data_diamond_numeric$Colour
x4<-data_diamond_numeric$Clarity
x5<-data_diamond_numeric$Cut.
x6<-data_diamond_numeric$Certification
x7<-data_diamond_numeric$Polish
x8<-data_diamond_numeric$Symmetry

model1_diamond<-lm(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8)
model1_diamond
summary(model1_diamond)
```

```{r}
predict_function_diamond<-function(x1, x2, x3, x4, x5, x6, x7, x8) {
  return (2075.38 + 1956.52 * x1 - 725.94 * x2 - 36.48 * x3 + 32.82 * x4 + 23.65 * x5 - 46.48 * x6 +  22.77 * x7 + 23.64 * x8)
}
predict_function_diamond(0.9, 1, 2, 2, 4, 4, 2, 4)
```

* If we set the wholesaler as 1, then the quote priced is fairly. If the wholesaler > 1. then the diamond overpriced.

* If the Carat increase by 1, the price will increase by $1956.52.
* If the Wholesaler increase by 1 level, the price will decrease by $725.94.
* If the Colour increase by 1 level, the price will decrease by $36.48.
* If the Clarity increase by 1 level, the price will increase by $32.82.
* If the Cut increase by 1 level, the price will increase by $23.65.
* If the Certification increase by 1 level, the price will decrease by $46.48.
* If the Polish increase by 1 level, the price will increase by $22.77.
* If the Symmetry increase by 1 level, the price will increase by $23.64.

## 2.3 Model2

```{r}
model2_diamond<-lm(y ~ x1 + x3 + x4 + x5 + x6 + x7 + x8)
model2_diamond
summary(model2_diamond)
```

* If we drop the wholesale column, the R-squared decreased by 0.6. If we set the significance level as 0.01, the summary of the model shows that the p-value for wholesaler is less than the significance level.Therefore, I reject null hypothesis that the wholesaler is equal to 0 at 0.01 level of significance.

```{r}
predict_function_diamond<-function(x1, x3, x4, x5, x6, x7, x8) {
  return (2075.38 + 1956.52 * x1 - 36.48 * x3 + 32.82 * x4 + 23.65 * x5 - 46.48 * x6 +  22.77 * x7 + 23.64 * x8)
}
predict_function_diamond(0.9, 2, 2, 4, 4, 2, 4)
```
* Based on the summaries of two models, I would not say that the model2 is better than the model1. 95.74% of the data fit the regression model1, and 89.73% of the data fit the regression model2.


